---
title: "Foodborne module"
output:
  pdf_document:
    citation_package: natbib
    keep_tex: true
bibliography: reference.bib
---

This document presents an informal documentation of the foodborne module based 
on the studies @collineau20 and @faverjon22. The foodborne module consists of 
four submodules that model different stages of broiler production, namely:

- Production submodule
  - Flock stage
  - Transport stage
- Processing submodule
  - Scalding stage
  - Defeathering stage
  - Evisceration stage
  - Washing stage
  - Chilling stage
  - Portioning stage
- Post-processing submodule
  - Retail stage
  - Consumer transport stage
  - Home storage stage
- Home preparation submodule

At the end of each submodule (stage) the module simulates 1) prevalence of 
infected birds in a flock and 2) the total number of bacterias on the carcasses. 

## Production submodule

The production submodule is divided into two stages namely, the flock stage and
the transport stage. The same framework is followed as proposed by @collineau20 
with the following modification in parameter values.

```{r, include=FALSE}
# Load the knitr package
library(knitr)

# Production submodule
production_inputs <- data.frame(
  Type = c(c("Input", "Input"), rep("Parameter", 9)),
  Symbol   = c(
    "C_barn",
    "Prev_wfp_col_base",
    "Prev_Farm_type",
    "Prev_wfp_ext.min",
    "Prev_wfp_ext.max",
    "F_cross_trans",
    "F_transp",
    "C_prod_n",
    "Amount_fec",
    "N_transp",
    "N_contact"
  ),
  Value    = c(
    "Farm module",
    "Farm module",
    "Beta$(26, 164)$",
    "$0.03$",
    "$0.5$",
    "U$(0, 0.5)$",
    "$1.41$",
    "Tri$(4.4,4.8,4.6)$",
    "Tri$(1,50,10)$",
    "U$(0,4)$",
    "Pert$(1.5, 4.5, 3)$"
  ),
  Source   = c(
    "Sarnino et al. (in prep.)",
    "Sarnino et al. (in prep.)",
    "Faverjon et al. (2022)",
    "Collineau et al. (2020)",
    "Collineau et al.(2020)", 
    "Collineau et al.(2020)",
    "Collineau et al. (2020)",
    "Faverjon et al. (2022)",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)"
  )
)

# production submodule dataframe

production_calcul <- data.frame(
  Type = c(rep("Variable", 8), rep("Output", 4)),
  Symbol = c("Prev_f", 
             "B_flock_status", 
             "Prev_wfp_col", 
             "C_btp", 
             "Prev_wfp_ext", 
             "P_ccbf",
             "P_ccwf",
             "P_pos",
             "Prev_prod_p",
             "Prev_prod_n",
             "C_prod_p",
             "C_prod_n"),
  Value = c("Prev_Farm_type",
            "rbinom(1, Prev_f)",
            "Prev_wfp_col_base",
            "C_barn * Amount_fec",
            "Pert(Prev_wfp_ext.min, Prev_wfp_col, Prev_wfp_ext.max)",
            "(1 - (1 - Prev_f)^N_transp) * F_cross_trans",
            "1 - (1- Prev_wfp_ext)^N_contact",
            "P_ccwf + P_ccbf - (P_ccwf * P_ccbf)",
            "Prev_wfp_ext + (1- Prev_wfp_ext) * P_pos",
            "P_ccbf",
            "C_btp * F_transp",
            "C_prod_n")
)
```
```{r}
kable(production_inputs, caption = "Production module inputs")
kable(production_calcul, caption = "Production module variables")
```
## Processing submodule

The processing module is divided in six stages namely, scalding, defeathering,
evisceration, washing, chilling and portioning. At each stage the prevalence and
number of bacterias are updated and used as inputs for the next stage/submodule.

Inputs: C_prod_p, Prev_prod_p, C_prod_n, Prev_prod_n

Outputs: C_proc_p, Prev_proc_p, C_proc_n, Prev_proc_n

### Scalding stage:

Following the same principle as @collineau20, only soft scalding is considered 
with reduction rate F_scald_soft from @althaus17 (@faverjon22):
```{r, include=FALSE}
# Production submodule
scalding <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 3), rep("Variable", 2), 
               rep("Output", 4)),
  Symbol   = c(
    "Prev_prod_p",
    "Prev_prod_n",
    "C_prod_p",
    "C_prod_n",
    "C_prod.min",
    "C_prod.max",
    "F_scald_soft",
    "Prop_scald",
    "Th_scald",
    "Prev_scald_p",
    "Prev_scald_n",
    "C_scald_p",
    "C_scald_n"
  ),
  Value    = c(
    "Prod. module",
    "Prod. module",
    "Prod. module",
    "Prod. module",
    "$5.5$",
    "$6.5$",
    "N$(4.16, 1.08)$",
    "10^(-F_scald_soft)",
    "U(1, 10^C_prod.min, 10^C_prod.max)",
    "ifelse(C_scald_p$>$Th_scald, 1, Prev_prod_p)",
    "ifelse(C_scald_n$>$Th_scald, 1, Prev_prod_n)",
    "C_prod_p * Prop_scald",
    "C_prod_p * Prop_scald"
  )
)
```
```{r}
kable(scalding, caption = "Processing module: scalding stage") 
```
### Defeathering stage:

Same as @collineau20.

```{r, include=FALSE}
defeathering <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 5), rep("Output", 4)),
  Symbol   = c(
    "Prev_scald_p",
    "Prev_scald_n",
    "C_scald_p",
    "C_scald_n",
    "a_df",
    "b_df",
    "F_df",
    "P_cross_df_p",
    "P_cross_df_n",
    "Prev_scald_p",
    "Prev_scald_n",
    "C_df_p",
    "C_df_n"
  ),
  Value    = c(
    "Scalding",
    "Scalding",
    "Scalding",
    "Scalding",
    "$-1.91$",
    "$-1.95$",
    "Tri$(-4.01, -2.45, -0.89)$",
    "0.6",
    "N$(0.02, 0.000557) [0,1]$",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)"
  )
)

```
```{r}
kable(defeathering, caption = "Processing module: defeathering stage") 
```
### Evisceration stage:

Same as @collineau20 with updated parameters, namely, C_caecal from @robe19 and 
F_ev from @belluco16 (@faverjon22).

```{r, include=FALSE}
evisceration <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 5), rep("Output", 4)),
  Symbol   = c(
    "Prev_df_p",
    "Prev_df_n",
    "C_df_p",
    "C_df_n",
    "C_caecal",
    "P_cross_ev_p",
    "F_ev",
    "P_cut_vis",
    "spill_weight",
    "Prev_ev_p",
    "Prev_ev_n",
    "C_ev_p",
    "C_ev_n"
  ),
  Value    = c(
    "defeathering",
    "defeathering",
    "defeathering",
    "defeathering",
    "Tri$(0.54, 8.83, 8.69)$",
    "Tri$(0.03,	0.15,	0.07)$",
    "Tri$(0.06,	0.12,	0.091)$",
    "Tri$(0.14,	0.23,	0.18)$",
    "U$(1, 10)$",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)",
    "Collineau et al. (2020)"
  )
)

```
```{r}
kable(evisceration, caption = "Processing module: evisceration stage") 
```
### Washing stage:

Following the discussion with Fédération des Industries Avicoles (FIA), two types 
of washing are implemented, namely, Inside Outside Bird Wash (IOBW) and IOBW along 
with Post Evisceration (PE) wash. About $50\%-75\%$ of the farms practice two types
of washing which translates to the indicator variable wash_type. Parameters are 
taken from @collineau20.

```{r, include=FALSE}
washing <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 4), rep("Variable", 3), rep("Output", 4)),
  Symbol   = c(
    "Prev_ev_p",
    "Prev_ev_n",
    "C_ev_p",
    "C_ev_n",
    "wash_type",
    "F_wash_IOBW",
    "F_wash_PE",
    "F_wash_adj",
    "wash_type",
    "Prop_wash_init",
    "Prop_wash",
    "Prev_wash_p",
    "Prev_wash_n",
    "C_wash_p",
    "C_wash_n"
  ),
  Value    = c(
    "evisceration",
    "evisceration",
    "evisceration",
    "evisceration",
    "ifelse($Y_{U[0, 1]}>0.625$, IOBW, IOBW+PE)",
    "Tri$(-1.29,	-0.92,	-0.55)$",
    "Tri$(-1.28,	-1.17,	-1.06)$",
    "U$(0,	0.5)$",
    "U$(0, 1)$",
    "ifelse(wash_type$>0.625$, 10^F_wash_IOBW, 10^(F_wash_IOBW+F_wash_PE))",
    "Prop_wash_init + (1 + F_wash_adj)",
    "Prev_ev_p",
    "Prev_ev_n",
    "C_ev_p * Prop_wash",
    "C_ev_n * Prop_wash"
  )
)
```
```{r}
kable(washing, caption = "Processing module: washing stage") 
```
### Chilling stage:

Following the discussion with Fédération des Industries Avicoles (FIA), two types 
of chilling are identified, namely, Air Chilling (AC) and Water Immersion Chilling 
(WIC). However the later is practiced for less than $2\%$ of the producers and thus
not implemented in the model. Moreover AC does not have the chance of cross 
contamination unlike WIC. Parameters are taken from @collineau20.

```{r, include=FALSE}
chilling <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 1), rep("Output", 4)),
  Symbol   = c(
    "Prev_wash_p",
    "Prev_wash_n",
    "C_wash_p",
    "C_wash_n",
    "Prop_chill",
    "Prev_chill_p",
    "Prev_chill_n",
    "C_chill_p",
    "C_chill_n"
  ),
  Value    = c(
    "washing",
    "washing",
    "washing",
    "washing",
    "Pert$(10^{0.308}/10^{1.6}, 10^{0.7}/10^{1.6}, 10^{1.092}/10^{1.6})$",
    "Prev_wash_p",
    "Prev_wash_n",
    "C_wash_p * Prop_chill",
    "C_wash_n * Prop_chill"
  )
)
```
```{r}
kable(chilling, caption = "Processing module: chilling stage") 
```
### Portioning stage:

Same as @collineau20 with the product type, namely, chicken breasts or whole 
carcasses is determined by an additional parameter Prop_product (see, e.g., 
@faverjon22) signifying the fraction of the raw product that represents the final 
processed product.

```{r, include=FALSE}
portioning <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 3), rep("Output", 4)),
  Symbol   = c(
    "Prev_chill_p",
    "Prev_chill_n",
    "C_chill_p",
    "C_chill_n",
    "Prop_cm",
    "P_skin",
    "Prop_product",
    "Prev_proc_p",
    "Prev_proc_n",
    "C_proc_p",
    "C_proc_n"
  ),
  Value    = c(
    "chilling",
    "chilling",
    "chilling",
    "chilling",
    "Pert$(0.01, 0.02, 1)$",
    "Beta$(1, 3.15)$",
    "0.25",
    "Prev_chill_p",
    "Prev_chill_n",
    "ifelse(Prop_product==1, C_chill_p, C_chill_p * P_skin * Prop_cm * Prop_product)",
    "ifelse(Prop_product==1, C_chill_n, C_chill_n * P_skin * Prop_cm * Prop_product)"
  )
)
```
```{r}
kable(portioning, caption = "Processing module: portioning stage") 
```

## Post-processing submodule
  
The post-processing sublmodule is divided into three main stages, namely, retail 
stage, consumer transport stage and home storage stage. The bacteria load changes
throughout these stages but the prevalence reamins the same. The final output 
prevalence at the end of the post-processing step is adjusted w.r.t the final load
(see, e.g., @collineau20).The effect os packaging on the bacteria load is also taken
into account, based on @faverjon22 and discussions with FIA. The module parameters 
related to bacteria growth are taken from @faverjon22 and the basic framework is
adapted from @collineau20.

Inputs: C_proc_p, Prev_proc_p, C_proc_n, Prev_proc_n

Outputs: C_pproc_p, Prev_pproc_p, C_pproc_n, Prev_pproc_n

```{r, include=FALSE}
pprocessing <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 19)),
  Symbol   = c(
    "Prev_proc_p",
    "Prev_proc_n",
    "C_proc_p",
    "C_proc_n",
    "T_growth_min",
    "T_growth_opt",
    "F_pack",
    "F_freezer",
    "Time_gen_min",
    "C_MPD",
    "Carcass_size_mean",
    "Carcass_size_sd",
    "Breast_size_mean",
    "Breast_size_sd",
    "Prop_carc_yield",
    "Time_retail",
    "Time_trans",
    "Time_fridge",
    "T_retail",
    "T_post_trans",
    "T_fridge",
    "Prop_breast_MAP",
    "Prop_carcass_fresh"
  ),
  Value    = c(
    "processing",
    "processing",
    "processing",
    "processing",
    "$7$",
    "Pert$(35, 40, 37)$",
    "Tri$(-0.3,	-0.1,	-0.2)$",
    "$0.33$",
    "$0.47$",
    "123000",
    "1495.3",
    "303.4",
    "111.6",
    "71.9",
    "U$(0.6, 0.65)$",
    "Tri$(1, 7, 3)$",
    "N$(69.6, 0.438)$",
    "N$(2.2, 0.0203)$",
    "Laplace$(3.3333, 2.6623, [-6.67, 19.44])$",
    "Sftd.Logistic$(-22.915, 29.371, 16.763, [-5.56, 20])$",
    "Laplace$(5.3, 2.2965, [-4.44, 16.11])$",
    "$0.75$",
    "$0.85$"
  )
)

pprocessing_calcul <- data.frame(
  Type = c(rep("Variables", 16), rep("Output", 4)),
  Symbol   = c(
    "breast_size",
    "carcass_size_init",
    "carcass_size",
    "Size",
    "C_max",
    "T_avg_trans",
    "Time_gen_retail",
    "Time_gen_trans",
    "Time_gen_fridge",
    "G_retail",
    "G_trans",
    "G_fridge",
    "C_ret",
    "Carcass_fresh",
    "Breast_MAP",
    "C_pack",
    "C_pproc_p",
    "C_pproc_n",
    "Prev_pproc_p",
    "Prev_pproc_n"
  ),
  Value    = c(
    "Gamma((mean / sd)^2, sd^2 / mean",
    "Gamma((mean / sd)^2, sd^2 / mean",
    "Prop_carc_yield * carcass_size_init",
    "ifelse(Prop_product == 1, carcass_size, breast_size)",
    "C_MPD * Size",
    "(T_retail + T_post_trans)/2",
    "(2) in Collineau et al. (2020)",
    "(2) in Collineau et al. (2020)",
    "(2) in Collineau et al. (2020)",
    "(1) in Collineau et al. (2020)",
    "(1) in Collineau et al. (2020)",
    "(1) in Collineau et al. (2020)",
    "C_proc_i * G_retail * G_trans * G_fridge",
    "Bin(Runs, Prop_carcass_fresh)",
    "Bin(Runs, Prop_breast_MAP)",
    "Based on FIA, Faverjon et al. (2022) and Collineau et al. (2020) on C_ret",
    "(4) in Collineau et al. (2020) on C_pack",
    "(4) in Collineau et al. (2020) on C_pack",
    "(3) in Collineau et al. (2020) on C_pproc_p and Prev_proc_p",
    "(3) in Collineau et al. (2020) on C_pproc_n and Prev_proc_n"
  )
)
```
```{r}
kable(pprocessing, caption = "Post-processing module: inputs and parameters") 
kable(pprocessing_calcul, caption = "Post-processing module: variables and outputs") 
```
## Home preparation submodule

Based on @collineau20

```{r, include=FALSE}
homeprep <- data.frame(
  Type = c(rep("Input", 4), rep("Parameter", 10), rep("Variable", 7), rep("Output", 8)),
  Symbol   = c(
    "Prev_pproc_p",
    "Prev_pproc_n",
    "C_pproc_p",
    "C_pproc_n",
    "Prop_protec",
    "Prop_fluid",
    "P_undercook",
    "V_dil_car",
    "D_value_alpha",
    "D_value_beta",
    "Time_protec",
    "T_protec",
    "P_h_wash",
    "V_ing",
    "V_dil",
    "R_ref",
    "tsf",
    "P_cc_p",
    "P_cc_n",
    "C_cc_p",
    "C_cc_n",
    "Prev_home_cook_p",
    "Prev_home_cook_n",
    "P_home_cc_p",
    "P_home_cc_n",
    "C_home_cook_p",
    "C_home_cook_n",
    "C_home_cc_p",
    "C_home_cc_n"
  ),
  Value    = c(
    "Post-proc. module",
    "Post-proc. module",
    "Post-proc. module",
    "Post-proc. module",
    "Pert$(0.1, 0.2, 0.16)$",
    "U$(0.01, 0.1)$",
    "N$(0.4, 0.00212)$",
    "U$(150, 250)$",
    "$8.58$",
    "$-0.139$",
    "Pert$(0.5, 1.5, 1)$",
    "Pert$(60, 65, 64)$",
    "U$(0.38, 1)$",
    "U$(0.5, 1.5)$",
    "Collineau et al. (2020)",
    "10^(D_value_beta * T_protec + D_value_alpha)",
    "Prop_fluid * V_ing / V_dil",
    "P_h_wash * Prev_pproc_p",
    "P_h_wash * Prev_pproc_n",
    "tsf * C_pproc_p",
    "tsf * C_pproc_n",
    "P_undercook * Prev_pproc_p",
    "P_undercook * Prev_pproc_n",
    "P_cc_p * (1- dpois(0, C_cc_p))",
    "P_cc_n * (1- dpois(0, C_cc_n))",
    "Prop_protec * 10^(log(C_pproc_p) - Time_protec / R_ref)",
    "Prop_protec * 10^(log(C_pproc_n) - Time_protec / R_ref)",
    "ifelse(C_cc_p <1, 1, C_cc_p)",
    "ifelse(C_cc_n <1, 1, C_cc_n)"
  )
)
```
```{r}
kable(homeprep, caption = "Home preparation module") 
```






