---
title: "R Notebook"
output: html_notebook
---

```{r, include=FALSE}
source(here::here("foodborne-module/load_libraries.R"))
source(here::here("foodborne-module/utilities/estimate_variables.R"))
source(here::here("foodborne-module/utilities/visualization.R"))
source(here::here("farm-module/run_farm_module_parallel.R"))

## Initialization

Runs <- 1000 # number of simulation to be performed

# Create dataframe with specified number of rows and column names
data <- data.frame(matrix(1:Runs, nrow = Runs, ncol = 1))
colnames(data) <- "Runs"
output <- data

# Load estimated variables metadata
input <- read_excel(here("foodborne-module/data-input/estimated_variables.xlsx"))

# Simulation of estimated variables
data <- estimate_variables(data, input, N=Runs)

# Run farm module to get initial load and prevalence
parallel_output <- batch_simulator_parallel(n_sim = Runs)

data$C_barn <- parallel_output[36, 1,]
data$Prev_wfp_col_base <- parallel_output[36, 2,]

# Risk calculation for each module
source(here::here("foodborne-module/Module_production.R"))
source(here::here("foodborne-module/Module_processing.R"))
source(here::here("foodborne-module/Module_postprocessing.R"))
source(here::here("foodborne-module/Module_homepreparation.R"))
```
## Concentration of ESBL *E. coli*
```{r, include=TRUE, warning=FALSE}
# Plot outputs
plot_load(output, plot_all = FALSE, flock = FALSE)
plot_load(output, plot_all = FALSE)
plot_load(output, plot_all = TRUE)
```
- These figures demonstrate the evolution of average bacteria load or the mean 
number of ESBL producing *E. coli* on broiler carcasses/portinos (upto/afterwards the 
chilling stage). 
- $1000$ independent Monte Carlo runs of the QMRA model were performed and the 
confidence intervals are based on $2.5 -97.5 \%$ empiricalquantiles.
- The figures show high skewness of the final concentration $C_{\rm home}$ after
home cooking.

### Summary statistics of concentration $C_{\rm home}$
```{r, include=FALSE}
# Filter outputs
output_p <- output %>%
  filter(B_flock_status == "p")

output_p_1plus <- output %>%
  filter(B_flock_status == "p" & C_home_cook > 1)

output_n <- output %>%
  filter(B_flock_status == "n")

output_n_1plus <- output %>%
  filter(B_flock_status == "n" & C_home_cook > 1)


load_df <- data.frame(
  Count = c(
    length(output$C_home_cook),
    length(output_p$C_home_cook),
    length(output_p_1plus$C_home_cook),
    length(output_n$C_home_cook),
    length(output_n_1plus$C_home_cook)
  ),
  Mean = c(
    mean(output$C_home_cook),
    mean(output_p$C_home_cook),
    mean(output_p_1plus$C_home_cook),
    mean(output_n$C_home_cook),
    mean(output_n_1plus$C_home_cook)
  ),
  Median = c(
    median(output$C_home_cook),
    median(output_p$C_home_cook),
    median(output_p_1plus$C_home_cook),
    median(output_n$C_home_cook),
    median(output_n_1plus$C_home_cook)
  ),
  Q5 = c(
    quantile(output$C_home_cook, 0.05),
    quantile(output_p$C_home_cook, 0.05),
    quantile(output_p_1plus$C_home_cook, 0.05),
    quantile(output_n$C_home_cook, 0.05),
    quantile(output_n_1plus$C_home_cook, 0.05)
  ),
  Q95 = c(
    quantile(output$C_home_cook, 0.95),
    quantile(output_p$C_home_cook, 0.95),
    quantile(output_p_1plus$C_home_cook, 0.95),
    quantile(output_n$C_home_cook, 0.95),
    quantile(output_n_1plus$C_home_cook, 0.95)
  ),
  Max = c(
    max(output$C_home_cook),
    max(output_p$C_home_cook),
    max(output_p_1plus$C_home_cook),
    max(output_n$C_home_cook),
    max(output_n_1plus$C_home_cook)
  )
)

rowname_list <- c("All", "+ flock", "+ flock, C_home > 1", "- flock", "- flock, C_home > 1")
rownames(load_df) <- rowname_list
```

```{r, include=TRUE}
library(knitr)
kable(load_df, caption = "Final concentration after home cooking")
```
- The table shows summary statistics for different classes based on initial flock 
status (positive/negative) or higher $C_{\rm home} > 1$ values.
- Significant difference between the mean and the median is observed;

## Prevalence of contaminated broilers/portions
```{r, include=TRUE, warning=FALSE}
plot_prev(output, plot_all = FALSE, flock = FALSE)
plot_prev(output, plot_all = FALSE)
plot_prev(output, plot_all = TRUE)
```

- These figures demonstrate the evolution of flock prevalence among the broiler 
carcasses/portinos (upto/afterwards the chilling stage).
- The summary statistics of final prevalence $P_{\rm home}$ after home cooking, 
in chicken portions are comparable with results from FSA.

### Summary statistics of prevalence $P_{\rm home}$
```{r, include=FALSE}
prev_df <- data.frame(
  Count = c(length(output$Prev_home_cook),
            length(output_p$Prev_home_cook),
            length(output_n$Prev_home_cook)),
  Mean = c(format(mean(output$Prev_home_cook), digits = 2),
           format(mean(output_p$Prev_home_cook), digits = 2),
           format(mean(output_n$Prev_home_cook), digits = 2)),
  Median = c(format(median(output$Prev_home_cook), digits = 2),
             format(median(output_p$Prev_home_cook), digits = 2),
             format(median(output_n$Prev_home_cook), digits = 2)),
  Q5 = c(format(quantile(output$Prev_home_cook, 0.05), digits = 2),
         format(quantile(output_p$Prev_home_cook, 0.05), digits = 2),
         format(quantile(output_n$Prev_home_cook, 0.05), digits = 2)),
  Q95 = c(format(quantile(output$Prev_home_cook, 0.95), digits = 2),
          format(quantile(output_p$Prev_home_cook, 0.95), digits = 2),
          format(quantile(output_n$Prev_home_cook, 0.95), digits = 2)),
  Max = c(format(max(output$Prev_home_cook), digits = 2),
          format(max(output_p$Prev_home_cook), digits = 2),
          format(max(output_n$Prev_home_cook), digits = 2))
)

rownames(prev_df) <- c("All", "+ flock", "- flock")
```
```{r, include=TRUE}
kable(prev_df, caption = "Final prevalence after home cooking")
```
## Computation of risk

For the computation of risk we first define the following events:

- $E_{\rm carrier}$: A consumer gets the ESBL *E. coli* carriership from consuming
a chicken portion from a particular flock that is produced by a production 
process characterized by the set of input parameters $\theta$.
- $E_{\rm consum}$: A consumer consumes a contaminated chicken portion from a
particular batch characterized by $\theta$.

Further define $P_{\rm carrier}(c_{\rm home})$ that denotes the probability of 
getting the ESBL *E. coli* carriership from consuming a chicken portion 
contaminated with $c_{\rm home}$ bacterias,
$$
\begin{aligned}
P_{\rm carrier}(c_{\rm home}) &= P[E_{\rm carrier} \vert E_{\rm consum}, c_{\rm home}]\\
&= \mathbb{E}[\mathbf{1}_{E_{\rm carrier}} \vert \mathbf{1}_{E_{\rm consum}}=1, c_{\rm home}].\\
\end{aligned}
$$
This probability is modeled using a Beta-Poisson dose response model as proposed by Furusawa et al.(2024),
$$
P_{\rm carrier}(C_{\rm home}) = 1-(1+\frac{C_{\rm home}}{\omega})^{-\alpha}.
$$
Now we derive the conditional risk per portion denoted by 
$R^{\rm portion}(P_{\rm home}, C_{\rm home})$, which is the probability of getting
the ESBL *E. coli* carriership from chicken protion consumption, from a particular
flock characterized by $P_{\rm home}$ and $C_{\rm home}$.

$$
\begin{aligned}
R^{\rm portion}(p_{\rm home}, c_{\rm home}) &= P[E_{\rm carrier} \vert 
P_{\rm home} = p_{\rm home}, C_{\rm home} = c_{\rm home}]\\
&= \mathbb{E}[\mathbf{1}_{E_{\rm carrier}} \vert p_{\rm home}, c_{\rm home}] \\
&= \mathbb{E}\left[\mathbb{E}\left[\mathbf{1}_{E_{\rm carrier}} \vert \mathbf{1}_{E_{\rm consum}}, p_{\rm home}, c_{\rm home} \right] \vert p_{\rm home}, c_{\rm home}\right] \\
&= \mathbb{E}[\mathbf{1}_{E_{\rm carrier}} \vert \mathbf{1}_{E_{\rm consum}}=1, p_{\rm home}, c_{\rm home}] \cdot P[\mathbf{1}_{E_{\rm consum}}=1 \vert p_{\rm home}, c_{\rm home}]\\ &+ 0 \cdot P[\mathbf{1}_{E_{\rm consum}}=0 \vert p_{\rm home}, c_{\rm home}]\\
&= P_{\rm carrier}(c_{\rm home}) \cdot p_{\rm home} 
\end{aligned}
$$

Now the ultimate QoI, the risk $R^{\rm portion}_{\theta}$ per portion of getting 
the ESBL *E. coli* carriership from consumption of a chicken portion from a flock 
characterized by $\theta$ is given by,

$$
\begin{aligned}
R^{\rm portion}_{\theta} &= P[E_{\rm carrier}] \\
&= \mathbb{E}[\mathbf{1}_{E_{\rm carrier}}] \\
&= \mathbb{E}\left[\mathbb{E}\left[\mathbf{1}_{E_{\rm carrier}} \vert P_{\rm home}, C_{\rm home}\right]\right] \\
&= \mathbb{E}[R^{\rm portion}(P_{\rm home}, C_{\rm home})] \\
&= \int P_{\rm carrier}(c_{\rm home}) \cdot p_{\rm home} \cdot p(p_{\rm home}, c_{\rm home}) \cdot \rm{d}p_{\rm home}\rm{d}c_{\rm home} 
\end{aligned}
$$
$R^{\rm portion}_{\theta}$ can be estimated using simple Monte Carlo integration
method, as shown in the following table. 
```{r, include=FALSE}
risk_df <- data.frame(
  Count = c(
    length(output$prob_carrier * output$Prev_home_cook),
    length(output_p$prob_carrier * output_p$Prev_home_cook),
    length(output_p_1plus$prob_carrier * output_p_1plus$Prev_home_cook),
    length(output_n$prob_carrier * output_n$Prev_home_cook),
    length(output_n_1plus$prob_carrier * output_n_1plus$Prev_home_cook)
  ),
  Mean = c(
    format(mean(
      output$prob_carrier * output$Prev_home_cook
    ), digits = 2),
    format(
      mean(output_p$prob_carrier * output_p$Prev_home_cook),
      digits = 2
    ),
    format(
      mean(output_p_1plus$prob_carrier * output_p_1plus$Prev_home_cook),
      digits = 2
    ),
    format(
      mean(output_n$prob_carrier * output_n$Prev_home_cook),
      digits = 2
    ),
    format(
      mean(output_n_1plus$prob_carrier * output_n_1plus$Prev_home_cook),
      digits = 2
    )
  ),
  Median = c(
    format(median(
      output$prob_carrier * output$Prev_home_cook
    ), digits = 2),
    format(
      median(output_p$prob_carrier * output_p$Prev_home_cook),
      digits = 2
    ),
    format(
      median(output_p_1plus$prob_carrier * output_p_1plus$Prev_home_cook),
      digits = 2
    ),
    format(
      median(output_n$prob_carrier * output_n$Prev_home_cook),
      digits = 2
    ),
    format(
      median(output_n_1plus$prob_carrier * output_n_1plus$Prev_home_cook),
      digits = 2
    )
  ),
  Q5 = c(
    format(
      quantile(output$prob_carrier * output$Prev_home_cook, 0.05),
      digits = 2
    ),
    format(
      quantile(output_p$prob_carrier * output_p$Prev_home_cook, 0.05),
      digits = 2
    ),
    format(
      quantile(
        output_p_1plus$prob_carrier * output_p_1plus$Prev_home_cook,
        0.05
      ),
      digits = 2
    ),
    format(
      quantile(output_n$prob_carrier * output_n$Prev_home_cook, 0.05),
      digits = 2
    ),
    format(
      quantile(
        output_n_1plus$prob_carrier * output_n_1plus$Prev_home_cook,
        0.05
      ),
      digits = 2
    )
  ),
  Q95 = c(
    format(
      quantile(output$prob_carrier * output$Prev_home_cook, 0.95),
      digits = 2
    ),
    format(
      quantile(output_p$prob_carrier * output_p$Prev_home_cook, 0.95),
      digits = 2
    ),
    format(
      quantile(
        output_p_1plus$prob_carrier * output_p_1plus$Prev_home_cook,
        0.95
      ),
      digits = 2
    ),
    format(
      quantile(output_n$prob_carrier * output_n$Prev_home_cook, 0.95),
      digits = 2
    ),
    format(
      quantile(
        output_n_1plus$prob_carrier * output_n_1plus$Prev_home_cook,
        0.95
      ),
      digits = 2
    )
  ),
  Max = c(
    format(max(
      output$prob_carrier * output$Prev_home_cook
    ), digits = 2),
    format(max(
      output_p$prob_carrier * output_p$Prev_home_cook
    ), digits = 2),
    format(
      max(output_p_1plus$prob_carrier * output_p_1plus$Prev_home_cook),
      digits = 2
    ),
    format(max(
      output_n$prob_carrier * output_n$Prev_home_cook
    ), digits = 2),
    format(
      max(output_n_1plus$prob_carrier * output_n_1plus$Prev_home_cook),
      digits = 2
    )
  )
)

rownames(risk_df) <- rowname_list
```
```{r, include=TRUE}
kable(risk_df, caption = "Summary statistics of Risk")
```
However given the skewness of the distribution (as shown by the next figure) of 
$R^{\rm portion}(P_{\rm home}, C_{\rm home})$ more suitable measure of risk can
be envisaged, namely, super quantiles. 

```{r}
plot_histogram_and_ecdf(output$prob_carrier*output$Prev_home_cook, "cond. risk")
```



